### 2020-10-20 latex/vlp.tex
Updated VLP chapter label
Added more to the VLP proof about escaping compact sets. NEeds a lot of editing.

%/========== Variable Length Pendulum ==========/%
\chapter{Application of VNHCS: The Variable Length Pendulum}\label{sec:vlp}
\section{Motivation}
The variable length pendulum (VLP) is a classical underactuated dynamical system
which is often used to model the motion of a person on a swing
\cite{pumping_swing_standing_squatting,how_to_pump_a_swing}.
The VLP also represents the motion of the load at the end of a crane, 
the (simplified) motion of a gymnast on a bar
\cite{pendulum_length_giant_gymnastics}, and the tuned-mass-damper systems which
stabilize skyscrapers \cite{vlp_tuned_mass_damper}.

@@

@@

   Suppose the initial condition of the VLP's unactuated variables is
   \((q_0,p_0) \not \in \left\{(0,0),(\pm\pi,0)\right\}\).
   A VNHC of the form 

@@

   Let \(\Gamma\) be the constraint manifold for \(h\).

@@

   Let \(D = \Gamma \backslash \{(\pm \pi, 0)\}\)
   be the domain of interest on which we study this negative-time system.

@@

   sets on \(D\). 

@@

   \begin{align*}

@@

   \end{align*}

@@

   The last step of the proof is to show that (\ref{eqn:negative-time-vlp})
   asymptotically converges to \((0,0)\). 
   For this, we will use a Krasovskii-LaSalle argument
   \cite{krasovskii_lasalle}.

@@

   in the \((q,p)\) plane, which means the VLP is gaining energy on average.

>>>>

%/========== Variable Length Pendulum ==========/%
\chapter{Application of VNHCS: The Variable Length Pendulum}\label{ch:vlp}
\section{Motivation}
The variable length pendulum (VLP) is a classical underactuated dynamical system
which is often used to model the motion of a person on a swing
\cite{pumping_swing_standing_squatting,how_to_pump_a_swing}.
The VLP also represents the motion of the load at the end of a crane, 
the (simplified) motion of a gymnast on a bar
\cite{pendulum_length_giant_gymnastics}, and the tuned-mass-damper systems which
stabilize skyscrapers \cite{vlp_tuned_mass_damper}.

@@

Before stating our theorem, we must rigorously what we mean by ``energy
injection" for the VLP. 
First, we want the phase of the VLP to move away from the origin; 
second, we want the system to reach any desired momentum in finite time.
These are captured mathematically by the following definitions.

\begin{defn}\label{defn:energy-injection}
   Let \(\dot{x} = f(x)\) be a planar ODE defined on \(D \subset \R^2\) and that
   \(f(0) = (0,0)\).
   The ODE experiences \textit{energy injection on \(D\)} if
   \begin{enumerate}
      \item The origin is a repeller 
         (or equivalently, it is asymptotically stable in negative time).
      \item For all compact sets \(K \subset D\), there exists \(T > 0\) where,
         for all \(t > T\), \(x(t) \not \in K\).
   \end{enumerate}
\end{defn}

\begin{defn}
   We say that a regular VNHC \(h\) \textit{injects energy into the VLP} if the
   constrained dynamics experience energy injection everywhere on the
   constraint manifold, except for possibly a set of measure zero.
\end{defn}

@@

   A regular VNHC of the form 

@@

   Let \(\Gamma = \mathbb{S}^1 \times \R \) be the constraint manifold for \(h\).

@@

   We will first study the upright equilibrium \((\pm \pi, 0)\). 
   The linearization of (\ref{eqn:negative-time-vlp}) at this
   equilibrium is given by
      df_{(\pm \pi,0)} = \begin{bmatrix}
         0 & -\frac{1}{mL(\pi)^2} \\
         -m g L(\pi) & 0 \\
   which has eigenvalues at \(\lambda = \pm \frac{g}{L(\pi)}\).
   Since this equilibrium has only one negative eigenvalue, the stable manifold
   theorem tells us that the set of initial conditions on \(\Gamma\) which
   converge to \((\pm \pi,0)\) is one-dimensional \textbf{SOURCE?}.
   Label this set of initial conditions
      \Pi_0 := \left\{ (q_0,p_0) \mid \lim_{t \to \infty} (q(t),p(t)) = (\pm pi, 0) \right\}
   Since \(\Gamma\) is 2D, \(\Pi_0\) is a set of measure zero.
   We therefore assume the system is not initialized on \(\Pi_0\), and will show
   that the origin is almost everywherethat globally asymptotically stable.
   % TODO: show (0,0) is a repeller and that all compact sets are left, forever,
   % in finite time
%   Let \(D = \Gamma \backslash \{(\pm \pi, 0)\}\)
%   be the domain of interest on which we study this negative-time system.

@@

   sets on \(\Gamma\). 

@@

   \begin{align}\label{eqn:vlp-proof-neg-inv}

@@

   \end{align}

@@

   Define \(D := \Gamma \backslash \Pi_0\) to be the domain of viable initial
   conditions.
   Using a Krasovskii-LaSalle argument \cite{krasovskii_lasalle}, 
   we will prove that (\ref{eqn:negative-time-vlp}) asymptotically converges to
   \((0,0)\) if it is initialized on \(D\).

@@

   in the \((q,p)\) plane, which means the origin is a repeller.
   This proves the first property of Definition \ref{defn:energy-injection}
   holds for the positive-time system; we need only show the second property
   also holds to complete the proof.
   We will prove the second property by contradiction.
   Let \(x(t)\) be the orbit for the positive-time system and assume
   \(x(0) \in K\backslash \Pi_0\).
   Suppose there exists a compact set \(K \subset \Gamma\) so that, for all
   \(T > 0\), there exists \(t > T\) where \(x(t) \in K\).
   Let 
      k := \sup\left\{E_\text{avg}(q,p) \mid (q,p) \in K \right\}
   be the largest energy value of \(K\), and let
      E_k := \left\{ (q,p) \in \Gamma \mid E_\text{avg}(q,p) \leq k \right\}
   be the largest energy sublevel set surrounding \(K\).
   Note that \(E_k\) is compact and negatively invariant (it is positively
   invariant for the negative-time system).
   Because \(K \subset E_k\), it follows that \(x(t) \in E_k\). 
   Then, since \(E_k\) is negatively invariant, the half orbit
   \(x([0,t])\) is contained fully within \(E_k\).
   In particular, \(x(T) \in E_k\); 
   but \(T\) was chosen arbitrarily, which implies \(x(t) \in E_k\) for all
   \(t \in [0,\infty[\).
   Since \(E_k\) is compact, \(x(t)\) has a positive limit set in
   \(E_k\) which (by Poincar\'{e}-Bendixson) is either an equilibrium or a
   closed orbit \textbf{SOURCE?}.
   The system cannot have closed orbits, since this would violate asymptotic
   stability of \((0,0)\) in negative time. 
   Hence, the positive limit set must be one of the equilibria. 
   The origin is a repeller, which means \(x(t)\) must be converging to 
   \((\pm \pi,0)\).
   This is a contradiction, since we assumed \(x(0) \not \in \Pi_0\).
      \Rightarrow \Leftarrow
   We have shown that orbits of the constrained dynamics will leave compact sets in
   finite time, including sublevel sets of \(E_\text{avg}\).
   Thus, the VLP must be gaining energy because its momentum is increasing.
